\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-03-12 papers/methods/paper.tex
Anais feedback 1

Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}. Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB. For a recent review, see \citet{Huterer2018}. Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.

In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric confirmation. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae. Such increased statistical power demands increased fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.

As such, staggering effort is being put into developing more sophisticated supernovae analyses. First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, the role of simulations has become increasingly important in determining biases and validating models. \citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements. Instead, most techniques fit these observations (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics. As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$.

The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology. However, given some modelled underlying population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude. Specifically, for any given supernova our functional map may take the traditional form:
\begin{equation}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections}, \label{eq:standard}
\end{equation}
where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter. The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.

\subsection{Traditional Analyses}

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$. This function is given by
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
\end{equation}
where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review). The cosmological $\mu_C$ is defined as
\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \label{eq:mu0}\\
\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
\end{align}

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with two costs. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.

@@

To try and escape the limitations of the traditional analysis methodology, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions. \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and Flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}. In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.

@@

One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$). \citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, however evaluation of systematic uncertainty was only performed using the Tripp metric.

@@

Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies hierarchical Bayesian Models. With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to the light curve fitting for supernova. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well by correctly including in their likelihood the determinant of the global covariance matrix. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision. Their model was validated on thirty realisations of statistics.

Whilst there are a large amount of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.

@@

The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae type rates and redshift distributions, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.

@@

The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$. This can be thought of as the `true' values for the supernova, and with correct summary statistics and uncertainty would be normally distributed around the observed values such that 
\begin{equation}
P(\hat{m}_B,\hat{x}_1,\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m}_B,\hat{x}_1,\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
\end{equation}
However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian. We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation. As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. Each survey has its own set of $\kappa$ parameters. Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlled the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponded to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$.

@@

The underlying supernova population is characterised by underlying distributions in colour, stretch and magnitude. Analytic models often treat the colour and stretch populations with a skew normal and normal, respectively, and have them as independent. Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution. Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the model population possesses similar ability to change as a function of redshift.

@@

On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept simulation fits show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into either population as determined by their observed classification probability $\hat{p}$.

@@

It is now well known that host galaxy environment has a significant effect on supernova properties. The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses. The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right]
\end{equation}
We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes
\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot m. \label{eq:l2}
\end{align}

@@

This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses, however it is normally constrained to band systematics. That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, 10 SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. Denoting $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{align}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{align}

@@

As the model so far described in previous sections describe components of a basic likelihood $P(D|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find

>>>>

{\red Todo: get info on all used lowz surveys and introduce them properly. Help in terms of the right papers to cite would be appreciated.}

Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB. For a recent review, see \citet{Huterer2018}. Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.

In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power. The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples. The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae. Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.

As such, staggering effort is being put into developing more sophisticated supernovae analyses. The role of simulations has become increasingly important in determining biases and validating models. First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.  \citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases. Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}. Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either also use simulation corrections or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements. Instead, most techniques fit these observations (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics. As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$.

The underlying actual supernova population is not such a clear cut, and accurately characterising this population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology. However, given some modelled underlying population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude. Specifically, for any given supernova our functional map may take the traditional form:
\begin{equation}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections}, \label{eq:standard}
\end{equation}
where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter. The $\text{corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}. The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.

\subsection{Traditional Cosmology Analyses}

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$. This function is given by
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
\end{equation}
where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review). The cosmological $\mu_C$ is defined as
\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}\\
\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
\end{align}

The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly. The speed comes with several limitations. Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty. Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation. Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae. As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.

@@

To try and escape the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions. \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and Flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}. In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.

@@

One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$). \citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, although evaluation of systematic uncertainty was only performed using the Tripp metric.

@@

Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies hierarchical Bayesian Models. With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision. Their model was validated on thirty realisations of statistics.

Whilst there are a large amount of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.

@@

The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae rates and redshift distributions, and host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.

@@

The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$. This can be thought of as the `true' values for the supernova, and with correct summary statistics and uncertainty would be normally distributed around the observed values such that 
\begin{equation}
P(\hat{m}_B,\hat{x}_1,\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m}_B,\hat{x}_1,\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
\end{equation}
However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian. We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation. As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. Each survey has its own set of $\kappa$ parameters. Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$.

@@

The underlying supernova population is characterised by underlying distributions in colour, stretch and magnitude. Analytic models often treat the colour and stretch populations independently with a skew normal and normal, respectively. Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution. Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the model population possesses similar ability to change as a function of redshift.

@@

On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SN Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept analysis on simple simulation data show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into the SN Ia population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.

@@

It is now well known that host galaxy environment has a significant effect on supernova properties. The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses. The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right].
\end{equation}
We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes
\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot m. \label{eq:l2}
\end{align}

@@

This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses, however it is normally constrained to band systematics. That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. Denoting $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{align}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{align}

@@

As our model so far described in previous sections describe components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
